---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bayespred

<!-- badges: start -->
<!-- badges: end -->

`bayespred` implements the pragmatic Bayesian workflow for clinical prediction
modelling described in Sadatsafavi & Riley (2026). It fits logistic regression
under four shrinkage priors and produces **self-contained** model objects whose
`predict()` method requires only base R at deployment time.

## Installation

```r
# Install dependencies
install.packages(c("mgcv", "brglm2"))

# Install from GitHub
# install.packages("pak")
pak::pak("resplab/bayespred")
```

## Quick start

```{r example}
library(bayespred)

# Low birth weight study (Hosmer & Lemeshow 1989), shipped with R via MASS
data(birthwt, package = "MASS")
birthwt$race <- factor(birthwt$race, levels = 1:3,
                       labels = c("white", "black", "other"))

# Fit under log-F(2) prior (default)
fit <- bpmfit(low ~ age + lwt + race + smoke + ht + ui,
              data = birthwt, prior = log_f(m = 2))

# New patient
new_pt <- data.frame(age = 26, lwt = 130, race = "white",
                     smoke = 1, ht = 0, ui = 0)

# The predict() by default returns the posterior mean (using an accurate quadrature method)
predict(fit, new_pt)

# With 95% credible interval
predict(fit, new_pt, interval = 0.95)

# One can request the plug-in estimate (based on [penalised] maximum likelihood)
predict(fit, new_pt, method="pe") # pe stand for plug-in estimate

# Compare all four priors
priors <- list(flat = flat(), jeffreys = jeffreys(),
               logf = log_f(m = 2), ridge = bridge())
sapply(priors, function(p) {
  predict(bpmfit(low ~ age + lwt + race + smoke + ht + ui,
                 data = birthwt, prior = p), new_pt)
})
```

## PM projection

`bpmproject()` regresses the posterior-mean soft labels onto a design matrix,
producing a standalone deployable model with no covariance matrix. All three
aspects default to the main fit but can be overridden: `formula` (predictor
set), `family` (link function), and `data` (can be the development sample,
local site data, or any external dataset).

```{r bpmproject}
# Self-projection: same predictors as the main model
proj <- bpmproject(fit)
coef(proj)
predict(proj, new_pt)

# Custom projection: reduce to key predictors
proj_simple <- bpmproject(fit, formula = ~ lwt + smoke + ht, data = birthwt)
coef(proj_simple)
predict(proj_simple, new_pt)

# Linear probability model — useful for nomograms
proj_linear <- bpmproject(fit, family = gaussian(link = "identity"))
coef(proj_linear)
```

The object returned by `bpmproject()` is fully self-contained and requires only
base R at deployment time.


## Full posterior

`posterior()` returns a `bpm` object that contains the MAP estimate,
Laplace-approximated covariance, and link function. A `bpm` object is
self-contained and can be used with `predict()` directly.

```{r federated}
post <- posterior(fit)     # bpm object: MAP + posterior covariance + link

# predict() works directly on the bpm object for a new patient
predict(post, new_pt)
```

## Reference

Sadatsafavi M, Riley RD (2026). Progression to the mean: A practical Bayesian workflow for the development and deployment of clinical prediction models *[https://arxiv.org/abs/2605.19163]*.
