---
title: "Pragmatic Bayesian Prediction Modelling with BayesCPM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pragmatic Bayesian Prediction Modelling with BayesCPM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

`BayesCPM` implements the pragmatic Bayesian workflow for clinical prediction
modelling described in Sadatsafavi & Riley (2026). The key insight is that the
*posterior mean* (PM) predicted probability — rather than the plug-in estimate
(PE) — is the correct target at deployment time under a proper scoring rule.

This vignette illustrates the full workflow on a simulated dataset that mimics
a binary 30-day outcome study.

## Simulated dataset

```{r data}
library(bayespred)
set.seed(2026)
n <- 2000

dat <- data.frame(
  age    = rnorm(n, mean = 65, sd = 12),
  sbp    = rnorm(n, mean = 130, sd = 20),
  sex    = factor(sample(c("male", "female"), n, replace = TRUE)),
  killip = factor(sample(c("I", "II", "III"), n, replace = TRUE,
                         prob = c(0.7, 0.2, 0.1)))
)
lp <- -4 + 0.04 * dat$age - 0.02 * dat$sbp +
       0.3 * (dat$sex == "male") +
       0.8 * (dat$killip == "II") + 1.5 * (dat$killip == "III")
dat$outcome <- rbinom(n, 1L, plogis(lp))
cat(sprintf("n = %d,  events = %d (%.1f%%)\n",
            n, sum(dat$outcome), 100 * mean(dat$outcome)))
```

## Fitting under different priors

```{r fit-all}
form <- outcome ~ age + sbp + sex + killip

fit_flat  <- bpm(form, data = dat, prior = flat())
fit_jeff  <- bpm(form, data = dat, prior = jeffreys())
fit_logf  <- bpm(form, data = dat, prior = log_f(m = 2))
fit_ridge <- bpm(form, data = dat, prior = bridge())
```

Compare coefficient estimates across priors:

```{r coef-table}
coef_tab <- sapply(
  list(flat = fit_flat, jeffreys = fit_jeff, logf2 = fit_logf, ridge = fit_ridge),
  coef
)
round(coef_tab, 3)
```

## Prediction for a new patient

```{r new-patient}
sam <- data.frame(age = 72, sbp = 105, sex = "male", killip = "II")
```

### Plug-in estimate vs posterior mean

```{r point-preds}
pe <- predict(fit_logf, sam, method = "pe")
pm <- predict(fit_logf, sam, method = "pm")
cat(sprintf("PE = %.3f   PM = %.3f\n", pe, pm))
```

### Credible intervals

```{r cri}
predict(fit_logf, sam, method = "pm", interval = 0.95)
predict(fit_logf, sam, method = "pe", interval = 0.95)
```

The PM lies inside but not at the centre of the interval (Jensen's inequality).

### MacKay approximation

```{r mackay}
predict(fit_logf, sam, method = "pm_mackay")
```

## PM projection for simplified deployment equations

`project_pm()` regresses the posterior-mean soft labels back onto a design
matrix, producing a standalone `bpm_proj_pm` object with no covariance matrix.
It can use the same predictors as the main model (self-projection) or a simpler
custom formula.

```{r projpred}
proj <- project_pm(fit_logf)
round(coef(proj), 3)

predict(proj, sam)
```

The projected model is self-contained and can be saved and reloaded
independently of the original fit.

## Model summary

```{r summary}
summary(fit_logf)
```

## All four priors on the same patient

```{r compare-all}
sapply(
  list(flat = fit_flat, jeffreys = fit_jeff, logf2 = fit_logf, ridge = fit_ridge),
  predict,
  newdata = sam,
  method = "pm"
)
```

## Reference

Sadatsafavi M, Riley RD (2026). Progression to the mean: A practical Bayesian workflow for the development and deployment of clinical prediction models *[https://arxiv.org/abs/2605.19163]*.
